Application of a Cut-Cell Immersed Boundary Method for Wildfire Simulation over Complex Terrain

### 2020-11-01 Vanella_Article_Atmosphere_2020.tex
Update on Overleaf.

@@

Wildland fire modeling typically focuses on tracking the fire front, which can have a complex shape that moves and deforms based on local conditions like fuel loading, moisture content, wind, and terrain. The fire front requires a mathematical representation. The Lagrangian approach is based on a polygonal mesh of control markers that represents the front~\cite{Finney:FARSITE,Bova:IJWF2015}. On the other hand, an Eulerian representation allows for a fixed two-dimensional mesh on which the fire front is defined implicitly by a scalar field. This scalar field is called the level set function and numerical methods that solve for it in time are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}. Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,mcgratta_2013,LAUTENBERGER_2013}. A practical comparison of the two techniques as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in reference~\cite{Bova:IJWF2015}.

Because the spread rate of the fire front depends in part on the local wind conditions, a well resolved wind field over complex terrain should improve the accuracy of the overall model. The de facto physics-based technique for practical simulation of wind is Large Eddy Simulation (LES)~\cite{coen_2013,Mell:IJWF2007}. For fire simulation over complex terrain, a range of techniques are being developed, in particular, in the coupling of the fire and the wind. For fast simulation of outdoor fires, an immersed boundary method has been interfaced with a Level Set scheme~\cite{Bova:IJWF2015} and the Rothermel fire spread model. The wind speed input for Rothermel's model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution. These are in turn also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept. This is a first step at implementing a level set approach in a LES wind calculation with combustion. It is recognized that, in the empirically-based development of the Rothermel formula, the wind speed was not influenced by the local buoyancy induced flow.

@@

For fire modeling on the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}. Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\bar{p}(z,t)$ used on the equation of state (the ideal gas law is used), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.

Consider a mixture composed of $N$ chemical species $\alpha$, moving on a fixed point $\mathbf{x}$ in space with a mass weighted average velocity $\mathbf{u}(\mathbf{x},t)$. If $\rho(\mathbf{x},t)$ is the mixture density, the mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, where $\rho_\alpha(\mathbf{x},t)$ is the species mass density and $\rho = \sum \rho_\alpha$. The scalar transport and momentum equations take the form:
\begin{eqnarray}
   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \rho Y_\alpha  \mathbf{u} ) &=& - \nabla \cdot \mathbf{J_{d}}_\alpha + \dot{m}_\alpha'''  +
    \dot{m}_{b,\alpha}'''\; , \; \alpha=1,\dots,N \label{eqn:spectran} \\
    \frac{\partial \mathbf{u}}{\partial t} - \mathbf{u} \times \boldsymbol{\omega} + \nabla H - \tilde{p} \, \nabla \left( 1/\rho\right) &=&
    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau} \right] \label{eqn:momtran}
\end{eqnarray}
where $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ are the diffusive fluxes for component $\alpha$. Fick's Law for binary diffusion respect to a background species is assumed, $D_\alpha$ is the diffusivity of $\alpha$ respect to the background species.  To enforce realizability, $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$, we lump errors in diffusive transport into the most abundant species locally \cite{McDermott:2015}. The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification. In the momentum equation the head is $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is an equlibrium density.  The gravity vector is $\mathbf{g}=(0,0,g_z)$, $\mathbf{f}_{b}$ is the particle force term, and $\boldsymbol{\tau}$ is a tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model. This last model is employed by default in large-eddy simulation (LES) of outdoor fires. Details of the model equations and their derivation are found in \cite{mcgratta_2013tr}.

Conservation of energy is achieved by forcing flow field to obey the following thermodynamic divergence constraint:
\begin{eqnarray}
    ( \nabla \cdot \mathbf{u} )^{th} &=&
    \left[ \frac{1}{\rho c_p T} - \frac{1}{\bar{p}} \right]
    \frac{\partial \bar{p}}{\partial t} + \frac{w \rho_0 g_z}{\rho c_p T} \nonumber \\
    &+& \frac{1}{\rho c_p T} \left[ \dot{q}''' + \dot{q}_b''' - \nabla \cdot \dot{\mathbf{q}}'' - \mathbf{u} \cdot \nabla (\rho h_s) \right] \nonumber \\
    &+& \frac{1}{\rho} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right) \left[ \dot{m}_\alpha'''  + \dot{m}_{b,\alpha}''' - \nabla \cdot (\mathbf{J_{d}}_\alpha) - \mathbf{u} \cdot \nabla (\rho Y_\alpha) \right] \label{eq:divth}
\end{eqnarray}
where $h_s$ is the local sensible enthalpy, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $\dot{\mathbf{q}}''$ is the heat flux, sum of conduction, convection and radiation, and $T$ is the local temperature. The mixture specific heat at constant pressure and molecular weight are $c_p=\sum_{\alpha =1}^N{c_{p,\alpha} Y_\alpha}$ and $\overline{W}=\left(\sum_{\alpha =1}^N{Y_\alpha /W_\alpha} \right)^{-1}$ respectively, built from the individual species counterparts as defined. Finally, $w,\rho_0,g_z$ on the stratification term are respectively, the local vertical velocity, density for standard conditions (depends on vertical coordinate), and vertical gravity acceleration. This thermodynamic divergence is derived from and acts as a proxy for the sensible enthalpy evolution equation~\cite{mcdermo_2014}. Equation (\ref{eq:divth}) is derived by factoring the divergence from the sensible enthalpy equation and applying the ideal gas law.  Given a background pressure and a local mass density the local temperature is computed from the ideal gas law as $T = \bar{p} \overline{W} / (\rho R) $.

@@

In figure~\ref{Fig:figure_1}a, a terrain is represented by its surface triangulation within an FDS cartesian mesh. Some cartesian cells and faces belonging to this mesh are transversed by the terrain surface. The remainder polyhedra and polygons that lay on the gas side of this intersected cartesian entities are called cut-cells and cut-faces. They discretize the fluid domain in the region surrounding the solid terrain. A detail of this unstructured mesh is shown in figure~\ref{Fig:figure_1}b. The resulting cut-cells and cut-faces are of varied sizes and shapes, and situations where more than one cut-cell per cartesian cell are not uncommon. As described later, small cells pose a temporal stability constraint in explicit time integration schemes that needs to be addressed.
It is noted that a two level of refinement grid hierarchy emerges. The coarse level is defined by the Cartesian entities, whereas the fine level is defined by the cut-cell or unstructured counterparts. Methods that solve discrete model equations on these grids are called cut-cell or embedded boundary methods~\cite{berger_2016}. Reliably defining the cut-cell mesh geometric properties and topology is a complex problem on itself, omitted here in order to keep the discussion succinct.
\begin{figure}[ht]
   \centering
   \includegraphics[trim = 12mm 42mm 8mm 25mm, clip,width=1\linewidth]{./figures/SketchFig1.pdf}
   \put(-305,-5){(a)}
   \put(-95,-4){(b)}
   \caption{Sketch of regular Cartesian and cut-cell grids around terrain: (a) An unstructured cut-cell region is defined on the gas phase side around a terrain $\mathbf{T}$. (b) Detail showing a slice representative of cut-cells and regular gas cells, terrain triangulation and two refinement level interpretation.}
   \label{Fig:figure_1}
\end{figure}

@@

The equations for chemical species transport are discretized using the finite volume method (FV)~\cite{eymard_2000,leveque_2002}. See figure~\ref{Fig:figure_2}a. Integrating equation~\eqref{eqn:spectran} over a cut-cell $ii$ with control volume $\Omega_{ii}$:

@@

 \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} d \Omega + \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)
      } d \Omega  = -\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \left(  \mathbf{J_{d}}_\alpha  \right)  } d \Omega + \int_{\Omega_{ii}} {( \dot{m}_\alpha'''+\dot{m}_{b,\alpha}''') } d \Omega \label{eq:intconvdiff}

@@

  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} d \Omega & = & \frac{\partial}{\partial t} \int_{\Omega_{ii}} {\rho Y_\alpha} d \Omega
  = \frac{\partial \widetilde{\rho \: Y_\alpha }_{ii}}{\partial t} V_{ii} \\
  \int_{\Omega_{ii}} {( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''')} d \Omega & = & (\widetilde{ \dot{m}_\alpha''' }_{ii}+ \widetilde{\dot{m}_{b,\alpha}'''}_{ii}) V_{ii} \label{eq:intcons}
where $V_{ii}$ is the volume of cell $ii$ and tildes imply cell averages. In the following the notation is simplified by dropping tildes, observing that quantities are cell or face averaged. Consider the FV discretization of the diffusive term (eqn.~\eqref{eq:intconvdiff}) on cut-cell $ii$ of figure~\ref{Fig:figure_2}a:

@@

\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \left(  \mathbf{J_{d}}_\alpha  \right)  } d \Omega =
    \int_{\partial \Omega_{ii}} { \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right) \cdot \hat{\mathbf{n}}_{ii} } \: d \partial \Omega = \sum^{nf_c}_{k=1}

@@

The integral over the cut-cell volume has been transformed in an area integral on its $nf_c=5$ faces using the divergence theorem. As these $k$ faces with areas $A_k$ and normals out of the control volume $\hat{\mathbf{n}}_{ii,k}$ are planar by construction, the method for evaluation of their mean diffusive fluxes $\left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k$ will define the spatial accuracy of the discretization. A centroid to centroid (i.e. $\Delta x_{cc}$ in figure~\ref{Fig:figure_2}a) finite differences and linear interpolation are used to approximate $\boldsymbol{\nabla} Y_\alpha$ and $\rho D_\alpha$ for each face belonging to the gas phase. Also, a normal probe approach~\cite{balaras_2004} is employed to sample information from the fluid to define fluxes in boundary cut-faces.

@@

    \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)} d \Omega =
     \int_{\partial \Omega_{ii}} { \left( \rho Y_\alpha \mathbf{u} \right) \cdot \hat{\mathbf{n}}_{ii} } \: d \partial \Omega =
     \sum^{nf_c}_{k=1} \left( \rho Y_\alpha \mathbf{u} \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \: A_k \label{eq:discfvadvcc}

@@

where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux limited interpolation to the cut-face~\cite{mcgratta_2013}. In the cut-cell region we use Godunov interpolation for the advective term. As the time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of equations~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known. The FV counterpart of the thermodynamic divergence expression for cut-cell $ii$ is:

@@

    ( \nabla \cdot \mathbf{u} )_{ii}^{th} \; V_{ii} &=&

@@

    \sum_{k=1}^{nf_c} \dot{\mathbf{q}}''_{ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k

@@

    \sum_{k=1}^{nf_c} \mathbf{J}_{\alpha,ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k

@@

where the over line terms refer to flux limited interpolation of corresponding scalars and terms defined with subscript $ii$ refer to cell defined quantities. All heat and mass fluxes and scalars are assumed known. Also, the vertical velocity $w_{ii}$ is interpolated to the cut-cell centroid. Details of the FDS time integration scheme can be found in references~\cite{mcgratta_2013,mcdermo_2014}. For each RK2 substep, the species transport equations are advanced in all FDS cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region. A similar procedure is done for the thermodynamic divergence.

@@

As an explicit time integrator is used, in general there will arise cut-cells whose small-size will penalize severely the time step. For these we link their unknown number to larger surrounding cells, which in turn defines mathematically an evolution equation for the union of both linked cells. Cell linking in the context of momentum equations can be found in reference~\cite{kirk_2003}.
In FDS, the momentum equations and pressure Poisson equation are solved in the cartesian mesh. Therefore, an immersed boundary method is used to reconstruct velocities in the cut-cell region. A divergence integral equivalence argument is used to transfer the the divergence from cut-cells to under laying cartesian cells, and build the source of the Poisson equation for the pressure.

@@

Collecting advective, shear stress and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within equation~\ref{eqn:momtran}, the model momentum transport problem can be casted as~\cite{mcgratta_2013}:
%
\begin{eqnarray}
  \frac{\partial \mathbf{u}(\mathbf{x},t)}{\partial t} &=& - \left[ \mathbf{F}(\mathbf{u},\mathbf{x},t) + \boldsymbol{\nabla} H(\mathbf{x},t) \right ]\label{eq:LowMachMom} \\
         \nabla \cdot \mathbf{u} (\mathbf{x},t) & = & \left(\nabla \cdot \mathbf{u} \right)^{th} \label{eq:LowMachDiv}
\end{eqnarray}
%
where equation~\eqref{eq:LowMachMom} is the momentum equation, subject to a specified divergence field, provided by equation~\eqref{eq:divth}.
Boundary conditions are prescribed for $\mathbf{u}(\mathbf{x},t)$ on boundaries, including the immersed terrain.
%Classical fractional step methods for time integration of incompressible or low Mach flow are based on two operations: First, momentum transport to obtain intermediate velocities, and second, projection of velocities into a target divergence space.
The previous equations are advanced in time using a fractional step method. As illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$. Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\nabla \cdot \mathbf{u}^{n+1} = \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th}$ known
%
\begin{eqnarray}
  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \left[ \mathbf{F}^n +  \boldsymbol{\nabla} H^n \right] \label{eq:LowMachMomEu}\\
  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th} \label{eq:LowMachDivEu}
\end{eqnarray}
%
where $\mathbf{u}^{n+1}$ represents a numerical solution at time $t_{n+1}$. This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator. As the potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is assumed responsible of enforcing the divergence condition and used on the projection step. Taking the divergence of equation~\eqref{eq:LowMachMomEu} and considering  the constraint~\eqref{eq:LowMachDivEu}, the two steps of the method are
%
\begin{enumerate}
  \item Solve Poisson equation for $H^n$:

\begin{equation}
   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u}^{n+1} \right)^{th} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
\end{equation}

@@

A coniferous tree (e.g. pine, spruce, fir) can be represented in FDS as a collection of Lagrangian particles forming a large cone. The trunk and heavier branches are typically neglected.

>>>>

\newcommand{\dV}{{\rm d}V}
\newcommand{\dS}{{\rm d}S}

@@

Wildland fire modeling typically focuses on tracking the fire front, which can have a complex shape that moves and deforms based on local conditions like fuel loading, moisture content, wind, and terrain. The fire front requires a mathematical representation. The Lagrangian approach is based on a polygonal mesh of control markers that represents the front~\cite{Finney:FARSITE,Bova:IJWF2015}. On the other hand, an Eulerian representation allows for a fixed two-dimensional mesh on which the fire front is defined implicitly by a scalar field. This scalar field is called the level set function and numerical methods that compute the function are called Level Set Methods (LSM)~\cite{Sethian:1999,Osher:2006}. Some wildfire solvers that use LSM to track the fire front are described in references~\cite{coen_2013,Bova:IJWF2015,mcgratta_2013,LAUTENBERGER_2013}. A practical comparison of the two techniques as implemented in the Fire Dynamics Simulator (FDS)~\cite{Mell:IJWF2007} and FARSITE~\cite{Finney:FARSITE} is provided in reference~\cite{Bova:IJWF2015}.

Because the spread rate of the fire front depends in part on the local wind conditions, a well resolved wind field over complex terrain should improve the accuracy of the overall model. The de facto physics-based technique for practical simulation of wind is Large Eddy Simulation (LES)~\cite{Linn:2007,coen_2013,Mell:IJWF2007}. For fire simulation over complex terrain, a range of techniques are being developed, in particular, in the coupling of the fire and the wind. For fast simulation of outdoor fires, an immersed boundary method has been interfaced with a Level Set scheme~\cite{Bova:IJWF2015} and the Rothermel fire spread model. The wind speed input for Rothermel's model results from the local velocity over the terrain which is affected by fire buoyancy and flow evolution. These are in turn also influenced by wind atmospheric conditions, accounted for within simulations using a mean wind forcing concept. This is a first step at implementing a level set approach in an LES wind calculation with combustion. It is recognized that in the empirically-based development of the Rothermel formula the wind speed is not influenced by the local buoyancy induced flow.

@@

For fire modeling in the gas phase, FDS employs a low Mach approximation for thermally driven buoyant and stratified flows~\cite{Rehm:1}. Under this assumption the pressure field $p$ can be viewed as the summation of two components: a background pressure $\bar{p}(z,t)$ used in the equation of state (ideal gas law), and a hydrodynamic pressure $\tilde{p}(x,y,z,t)$ resulting from fluid motion.

Consider a mixture composed of $N$ chemical species $\alpha$, moving on a fixed point $\mathbf{x}$ in space with a mass weighted average velocity $\mathbf{u}(\mathbf{x},t)$. If $\rho(\mathbf{x},t)$ is the mixture density, the mass fraction for species $\alpha$ is $Y_\alpha = \rho_\alpha / \rho$, where $\rho_\alpha(\mathbf{x},t)$ is the species mass density and $\rho = \sum \rho_\alpha$. The scalar transport and momentum equations take the form:
\begin{eqnarray}
   \frac{\partial \rho Y_\alpha}{ \partial t} + \nabla \cdot ( \rho Y_\alpha  \mathbf{u} ) &=& - \nabla \cdot \mathbf{J_{d}}_\alpha + \dot{m}_\alpha'''  +
    \dot{m}_{b,\alpha}'''\; , \; \alpha=1,\dots,N \label{eqn:spectran} \\
    \frac{\partial \mathbf{u}}{\partial t} - \mathbf{u} \times \boldsymbol{\omega} + \nabla H - \tilde{p} \, \nabla \left( 1/\rho\right) &=&
    \frac{1}{\rho} \left[ (\rho-\rho_0) \mathbf{g} + \mathbf{f}_{b} + \nabla \cdot \boldsymbol{\tau} \right] \label{eqn:momtran}
\end{eqnarray}
where $\mathbf{J_{d}}_\alpha=- \rho D_\alpha \boldsymbol{\nabla} Y_\alpha$ is the diffusive flux for component $\alpha$. Fick's Law for binary diffusion with respect to a background species is assumed, and $D_\alpha$ is the diffusivity of $\alpha$ with respect to the background species.  To enforce realizability, $\sum_\alpha \mathbf{J_{d}}_\alpha = 0$, errors in diffusive transport are lumped into the most abundant species locally~\cite{McDermott:2015}. The volumetric combustion source term for species $\alpha$ is $\dot{m}_\alpha'''(\mathbf{x},t)$  and $\dot{m}_{b,\alpha}'''$ is the contribution to species $\alpha$ from subgrid particle gasification. In the momentum equation the term $H=|\mathbf{u}|^2/2 + \tilde{p}/\rho$, $\tilde{p}$ is the perturbation pressure, and $\rho_0(z)$ is an ambient density of the atmosphere.  The gravity vector is $\mathbf{g}=(0,0,g_z)$, $\mathbf{f}_{b}$ is the particle force term, and $\boldsymbol{\tau}$ is a tensor accounting for molecular and subgrid turbulent stresses parameterized by an effective eddy viscosity model. This last model is employed by default in LES of outdoor fires. Details of the model equations and their derivation are found in Ref.~\cite{mcgratta_2013tr}.

Conservation of energy is achieved by forcing the flow field to obey the following thermodynamic divergence constraint:
\begin{eqnarray}
    ( \nabla \cdot \mathbf{u} )^{\rm th} &=&
    \left[ \frac{1}{\rho c_p T} - \frac{1}{\bar{p}} \right]
    \frac{\partial \bar{p}}{\partial t} + \frac{w \rho_0 g_z}{\rho c_p T} \nonumber \\
    &+& \frac{1}{\rho c_p T} \left[ \dot{q}''' + \dot{q}_b''' - \nabla \cdot \dot{\mathbf{q}}'' - \mathbf{u} \cdot \nabla (\rho h_s) \right] \nonumber \\
    &+& \frac{1}{\rho} \sum_\alpha \left( \frac{\overline{W}}{W_\alpha} - \frac{h_{s,\alpha}}{c_p T} \right) \left[ \dot{m}_\alpha'''  + \dot{m}_{b,\alpha}''' - \nabla \cdot (\mathbf{J_{d}}_\alpha) - \mathbf{u} \cdot \nabla (\rho Y_\alpha) \right] \label{eq:divth}
\end{eqnarray}
where $h_s$ is the local sensible enthalpy, $\dot{q}''', \dot{q}_b'''$ are heat release rates due to combustion and particle gasification, $\dot{\mathbf{q}}''$ is the heat flux vector (conduction, convection and radiation), and $T$ is the local gas temperature. The mixture specific heat at constant pressure and molecular weight are $c_p=\sum_{\alpha =1}^N{c_{p,\alpha} Y_\alpha}$ and $\overline{W}=\left(\sum_{\alpha =1}^N{Y_\alpha /W_\alpha} \right)^{-1}$ respectively. This thermodynamic divergence is derived from and acts as a proxy for the sensible enthalpy evolution equation~\cite{mcdermo_2014}. Equation~(\ref{eq:divth}) is derived by factoring the divergence from the sensible enthalpy equation and applying the ideal gas law.  Given a background pressure and a local mass density, the local temperature is computed from the ideal gas law as $T = \bar{p} \overline{W} / (\rho R) $.

@@

In Fig.~\ref{Fig:figure_1}, the terrain is represented by its surface triangulation within an FDS Cartesian mesh. Some Cartesian cells and faces belonging to this mesh are transversed by the terrain surface. The remaining polyhedra and polygons that lie on the gas side of these intersected geometries are called cut-cells and cut-faces. They discretize the fluid domain in the region surrounding the solid surface. A detail of this unstructured mesh is shown in Fig.~\ref{Fig:figure_1}. The resulting cut-cells and cut-faces are of varied sizes and shapes, and situations where there are more than one cut-cell per Cartesian cell are common. As described later, small cells impose a temporal stability constraint on explicit time integration schemes. A two-level grid refinement hierarchy emerges. The coarse level is defined by the Cartesian entities, whereas the fine level is defined by the cut-cell or unstructured components. Methods that solve discrete model equations on these grids are called cut-cell or embedded boundary methods~\cite{berger_2016}. Reliably defining the cut-cell mesh geometric properties and topology is a complex problem in and of itself, but beyond the scope of this paper.
\begin{figure}[ht]
   \centering
   \includegraphics[trim = 12mm 42mm 8mm 25mm, clip,width=1\linewidth]{./figures/SketchFig1.pdf}
   \put(-305,-5){(a)}
   \put(-95,-4){(b)}
   \caption{Sketch of regular Cartesian and cut-cell grids around terrain: (a) An unstructured cut-cell region is defined on the gas phase side around a terrain $\mathbf{T}$. (b) Detail showing a slice representative of cut-cells and regular gas cells, terrain triangulation and two refinement level interpretation.}
   \label{Fig:figure_1}
\end{figure}

@@

The equations for chemical species transport are discretized using the finite volume method (FV)~\cite{eymard_2000,leveque_2002}, as outlined in the diagram of Fig.~\ref{Fig:figure_2}. Integrating Eq.~(\ref{eqn:spectran}) over a cut-cell $ii$ with control volume $\Omega_{ii}$:

@@

 \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV + \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)
      } \, \dV  = -\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot \mathbf{J_{d}}_\alpha   } \, \dV + \int_{\Omega_{ii}} { \left( \dot{m}_\alpha'''+\dot{m}_{b,\alpha}''' \right) } \, \dV \label{eq:intconvdiff}

@@

@@

  \int_{\Omega_{ii}} {\frac{\partial \rho Y_\alpha}{\partial t}} \, \dV \approx \frac{\partial \widetilde{\rho Y_\alpha }_{ii}}{\partial t} V_{ii} \quad ; \quad
  \int_{\Omega_{ii}} { \left( \dot{m}_\alpha''' +\dot{m}_{b,\alpha}''' \right)} \, \dV \approx \left(\widetilde{ \dot{m}_\alpha''' }_{ii}+ \widetilde{\dot{m}_{b,\alpha}'''}_{ii} \right) V_{ii} \label{eq:intcons}
where $V_{ii}$ is the volume of cell $ii$ and tildes imply cell averages. In the following the notation is simplified by dropping tildes, observing that quantities are cell or face averaged. Consider the FV discretization of the diffusive term of Eq.~(\ref{eq:intconvdiff}) on cut-cell $ii$ of Fig.~\ref{Fig:figure_2}:
\int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot   \mathbf{J_{d}}_\alpha   } \, \dV =
    \int_{\partial \Omega_{ii}} { \left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right) \cdot \hat{\mathbf{n}}_{ii} } \, \dS = \sum^{n_{\rm f}}_{k=1}

@@

The integral over the cut-cell volume has been transformed in an area integral on its $n_{\rm f}=5$ faces using the divergence theorem. As these $k$ faces with areas $A_k$ and outward normals $\hat{\mathbf{n}}_{ii,k}$ are planar by construction, the method for evaluation of their mean diffusive fluxes $\left( - \rho D_\alpha \boldsymbol{\nabla} Y_\alpha \right)_k$ will define the spatial accuracy of the discretization. Centroid to centroid (i.e. $\Delta x_{\rm cc}$ in Fig.~\ref{Fig:figure_2}) finite differences and linear interpolation are used to approximate $\boldsymbol{\nabla} Y_\alpha$ and $\rho D_\alpha$ for each face belonging to the gas phase. Also, a normal probe approach~\cite{balaras_2004} is employed to sample information from the fluid to define fluxes in boundary cut-faces.

@@

    \int_{\Omega_{ii}} { \boldsymbol{\nabla} \cdot  \left(  \rho Y_\alpha \mathbf{u} \right)} \, \dV =
     \int_{\partial \Omega_{ii}} { \left( \rho Y_\alpha \mathbf{u} \right) \cdot \hat{\mathbf{n}}_{ii} } \, \dS \approx
     \sum^{n_{\rm f}}_{k=1} \left( \rho Y_\alpha \mathbf{u} \right)_k \cdot \hat{\mathbf{n}}_{ii,k} \, A_k \label{eq:discfvadvcc}

@@

where the advective flux for face $k$ is $\left( \rho Y_\alpha \mathbf{u} \right)_k = \overline{[\rho Y_\alpha]}_k \mathbf{u}_k$, and the over bar in $\overline{[\rho Y_\alpha]}_k$ means it is a flux-limited interpolation to the cut-face~\cite{mcgratta_2013}. In the cut-cell region we use Godunov interpolation for the advective term. As the time integration scheme in FDS is an explicit Runge-Kutta (RK2) method, all variables in the right hand side of Eqs.~\eqref{eq:discfvdiffcc}-\eqref{eq:discfvadvcc} are assumed known. The FV counterpart of the thermodynamic divergence expression for cut-cell $ii$ is:

@@

    ( \nabla \cdot \mathbf{u} )_{ii}^{\rm th} \; V_{ii} &=&

@@

    \sum_{k=1}^{n_{\rm f}} \dot{\mathbf{q}}''_{ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k

@@

    \sum_{k=1}^{n_{\rm f}} \mathbf{J}_{\alpha,ii,k} \cdot \hat{\mathbf{n}}_{ii,k} \: A_k

@@

where the over line terms refer to flux limited interpolation of corresponding scalars, and terms defined with subscript $ii$ refer to cell defined quantities. All heat and mass fluxes and scalars are assumed known. Also, the vertical velocity $w_{ii}$ is interpolated to the cut-cell centroid. Details of the FDS time integration scheme can be found in references~\cite{mcgratta_2013,mcdermo_2014}. For each RK2 substep, the species transport equations are advanced in all FDS Cartesian cells, and then the solution (explicit fluxes and scalar densities) recomputed on the unstructured cut-cell region. A similar procedure is done for the thermodynamic divergence.

@@

As an explicit time integrator is used, in general there will arise cut-cells whose small-size will penalize severely the time step. These cut-cells are linked to larger surrounding cells. Cell linking in the context of momentum equations can be found in Ref.~\cite{kirk_2003}. In FDS, the momentum equations and pressure Poisson equation are solved on the Cartesian mesh. Therefore, an immersed boundary method is used to reconstruct velocities in the cut-cell region. A divergence integral equivalence argument is used to transfer the divergence from cut-cells to the underlying Cartesian cells in order to build the source term of the Poisson equation for the pressure.

@@

Collecting advective, shear stress and force terms in $\mathbf{F}(\mathbf{u},\mathbf{x},t)$ within Eq.~(\ref{eqn:momtran}), the model momentum transport problem can be written~\cite{mcgratta_2013}:
\begin{eqnarray}
  \frac{\partial \mathbf{u}(\mathbf{x},t)}{\partial t} &=& - \mathbf{F}(\mathbf{u},\mathbf{x},t) - \boldsymbol{\nabla} H(\mathbf{x},t)  \label{eq:LowMachMom} \\
         \nabla \cdot \mathbf{u} (\mathbf{x},t) & = & \left(\nabla \cdot \mathbf{u} \right)^{\rm th} \label{eq:LowMachDiv}
\end{eqnarray}
where Eq.~\eqref{eq:LowMachMom} is the momentum equation, subject to a specified divergence field, provided by Eq.~\eqref{eq:divth}.
Boundary conditions are prescribed for $\mathbf{u}(\mathbf{x},t)$ on boundaries, including the immersed terrain.

The previous equations are advanced in time using a fractional step method. As illustration, consider their Forward Euler (FE) update from $t_n$ to $t_{n+1}=t_n + \Delta t$. Given $ \mathbf{u}^n=\mathbf{u}(\mathbf{x},t_n)$, $\nabla \cdot \mathbf{u}^{n+1} = \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th}$ known
\begin{eqnarray}
  \frac{\mathbf{u}^{n+1}-\mathbf{u}^{n}}{\Delta t} &=& - \mathbf{F}^n -  \boldsymbol{\nabla} H^n \label{eq:LowMachMomEu}\\
  \nabla \cdot \mathbf{u}^{n+1} &=& \left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th} \label{eq:LowMachDivEu}
\end{eqnarray}
where $\mathbf{u}^{n+1}$ represents a numerical solution at time $t_{n+1}$. This discrete FE update corresponds to the first sub-step of the FDS explicit RK2 integrator. As the potential field $H(\mathbf{x},t)$ does not have a time evolution equation, it is assumed responsible for enforcing the divergence condition and used on the projection step. Taking the divergence of Eq.~\eqref{eq:LowMachMomEu} and considering  the constraint, Eq.~\eqref{eq:LowMachDivEu}, the two steps of the method are
\begin{enumerate}
  \item Solve Poisson equation for $H^n$:

\begin{equation}
   \nabla \cdot \boldsymbol{\nabla} H^n = - \left[ \frac{\left( \nabla \cdot \mathbf{u}^{n+1} \right)^{\rm th} - \nabla \cdot \mathbf{u}^{n}}{\Delta t} \right] - \nabla \cdot \mathbf{F}^n \label{it:FSPoisson}
\end{equation}

@@

A coniferous tree (e.g., pine, spruce, fir) can be represented in FDS as a collection of Lagrangian particles forming a large cone. The trunk and heavier branches are typically neglected.
